---
title: "visualization"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```


# Load the data

```{r}

df = read.csv("policekillings202010.csv")

```

# Let's visualize something!

So, in the master file we have looked at race, age, method of crime, as well as some state-specific metrics.

Instead of having to look at state by state, wouldn't it be much better if we had a map of the US we could map data onto?

If you, like myself, can't, at gun point (pun intended), say where 70% of US states are located, you'll greatly benefit from this.

Let's get into it.

```{r}

# This is the ggplot function for our desired map. 

#ggplot() + 
  #geom_polygon(data=fifty_states, aes(x=lon, y=lat, group = group),color="white", fill="grey10")

# Oh we need to load and install a the library to run this!

#devtools::install_github("wmurphyrd/fiftystater")

# Then let's load the library.

library(fiftystater)

# And load the data necessary for the function to run
data("fifty_states")

ggplot() + 
  geom_polygon(data=fifty_states, aes(x=long, y=lat, group = group),color="white", fill="grey10")

```

Voila!!

Let's map some stats onto it.

But wait, before we do that, we need to make sure our data has longitude and latitude information for each observation - so we can map onto the X and Y coordinates in our plot above!


```{r}

# I found this CSV online. Why work when you don't have to work!

longlat = read.csv("state-medal-count.csv")

# Now what we need to do is to merge two dataframes together. We do it using inbuilt function merge() that takes three arguments: merge(df1, df2, by = "column")

#That last one is important - it means we need two indentical columns in our dataframes to successfully merge it.

# For us this will be state names - CA, WA, etc...

# As you can see, in the longlat df, the variable is "state", whereas in df it's "State".
# R will treat them as different columns - so we need to rename one.

longlat = longlat %>% 
  rename(State = state) %>% 
  select(-(medals))

# Ready now!

df = merge(df, longlat, by = "State")

# Boom!
# As you can see, there are now about 800 observations fewer.
# That's because our CSV from the world wide web doesn't have South Dakota and West Virginia.
# But who cares about them!

```

We are ready moms spaghetti.

Let's the topography of killings in the US!

```{r}

# First, we need to prepare a dataframe that displays # of killings in each state.

df_states = df %>% 
  group_by(State, lon, lat) %>% 
  summarize(count = n())

ggplot() + 
  geom_polygon(data=fifty_states, aes(x=long, y=lat, group = group),color="white", fill="grey92") +
  # This is where we start inserting our data on top of the map
  geom_point(data = df_states, aes(x = lon, y = lat, size = count)) +
  theme_void()

```

What else can we visualize?
